WIDS 2024 Final Project

### 2024-01-12 Code/Full R Code.qmd

**Insert abstract of project, where we got our data, what our data looks like, and other information here**

@@

```{r}
summary(income_distribution)
```
In this table, we can see that there are a few NA values in the tables, and some of the variables have a noticable skew. After some basic data cleaning is done we will see if these values remain

```{r}
summary(minimum_wage)
```
Only one column has NA values in this data frame, and in that column only 15 out of 2862 observations are missing values. Again, we will see if these values remain after our data cleaning

@@

```{r}
# seeing what we have left
head(union_wages)
head(union_states)
head(house)
head(income_distribution)
head(minimum_wage)
```
```{r}
colSums(is.na(house))
colSums(is.na(income_distribution))
colSums(is.na(minimum_wage))
colSums(is.na(union_wages))
colSums(is.na(union_states))
```
There are no more NA values

@@

```{r}
head(full_state_data)
summary(full_state_data)
```
After combining the dataset, we can see that not all of the variables are tracked every year, causing there to be NA values. We will keep this in mind as we continue our data analysis.

@@

```{r}
head(full_ntl_data)
summary(full_ntl_data)
```
There are no NA values in our national data set.

@@

This graph can be used to draw a number of interesting inferences. First, there is a decently dark line at the bottom that indicates a number of states still have no minimum wage. This will be examined in greater detailed in the next graph. Next, we can see that there is a significant shift post 2014 of raising the minimum wage at the state level. Before that time almost no state had a adjusted minimum wage over $10. Now, there are numerous states that surpass this value. Final, we can see that the federal minimum wage has been slowly losing value due to inflation since it was last raised in 2009.

@@

CB_by_decade2
ggsave("../Documents/Graph_Images/CB_by_decade2.png", width=7.29, height=5.41)
rm(CB_by_decade2)
```
These two graphs show very interesting results. First, we can see how unions have become politicized in the last 40 years. In the 1980s, there was almost no correlation between the amount of workers in a collective bargaining agreement and what the house delegation looked like in that state. By the 2010s, this had changed. From the second graph, we can see that the North Central and South regions are the main drivers behind this, seen in the steepening of the regression slope.

## Modeling

t test between union and non union wages
```{r}
# 1980s
full_ntl_data_1980s <- full_ntl_data |>
  filter(decade == "1980s")
t.test(full_ntl_data_1980s$union_wage_2020_adj, full_ntl_data_1980s$nonunion_wage_2020_adj)

@@

Linear regression of percent collective bargaining and party state average
```{r}
decade_region <- list(
  c("1980s", "Northeast"),
  c("1980s", "West"),
  c("1980s", "North Central"),
  c("1980s", "South"),
  c("2010s", "Northeast"),
  c("2010s", "West"),
  c("2010s", "North Central"),
  c("2010s", "South")
)

@@

Graph of regression estimate by decade and year
```{r}
Regression_plot <- ggplot(CB_party_regression, aes(x = Decade, y = Estimate, fill = Region)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = sprintf("P=%.3f", P_Value)),
             position = position_dodge(width = 0.9),
             vjust = -0.2,
             size = 3) +
  theme_minimal() +
  labs(title = "Linear Regression Estimate by Decade",
       subtitle = "Model of Percent Collective Bargaining vs Proportion of House Delegation That is Democratic")

@@

## Conclusion

>>>>

@@

```{r}
summary(income_distribution)
```

In this table, we can see that there are a few NA values in the tables, and some of the variables have a noticable skew. After some basic data cleaning is done we will see if these values remain

```{r}
summary(minimum_wage)
```

Only one column has NA values in this data frame, and in that column only 15 out of 2862 observations are missing values. Again, we will see if these values remain after our data cleaning

@@

```{r}
# seeing what we have left
head(union_wages)
head(union_states)
head(house)
head(income_distribution)
head(minimum_wage)
```

```{r}
colSums(is.na(house))
colSums(is.na(income_distribution))
colSums(is.na(minimum_wage))
colSums(is.na(union_wages))
colSums(is.na(union_states))
```

There are no more NA values

@@

```{r}
head(full_state_data)
summary(full_state_data)
```

After combining the dataset, we can see that not all of the variables are tracked every year, causing there to be NA values. We will keep this in mind as we continue our data analysis.

@@

```{r}
head(full_ntl_data)
summary(full_ntl_data)
```

There are no NA values in our national data set.

@@

This graph can be used to draw a number of interesting inferences. First, there is a decently dark line at the bottom that indicates a number of states still have no minimum wage. This will be examined in greater detailed in the next graph. Next, we can see that there is a significant shift post 2014 of raising the minimum wage at the state level. Before that time almost no state had a adjusted minimum wage over \$10. Now, there are numerous states that surpass this value. Final, we can see that the federal minimum wage has been slowly losing value due to inflation since it was last raised in 2009.

@@

CB_by_decade2
ggsave("../Documents/Graph_Images/CB_by_decade2.png", width=7.29, height=5.41)
rm(CB_by_decade2)
```

These two graphs show very interesting results. First, we can see how unions have become politicized in the last 40 years. In the 1980s, there was almost no correlation between the amount of workers in a collective bargaining agreement and what the house delegation looked like in that state. By the 2010s, this had changed. From the second graph, we can see that the North Central and South regions are the main drivers behind this, seen in the steepening of the regression slope.

## 2.4 Modeling

### 2.4.1 T-test between union and non union wages

```{r}
# 1980s
full_ntl_data_1980s <- full_ntl_data |>
  filter(decade == "1980s")
t.test(full_ntl_data_1980s$union_wage_2020_adj, full_ntl_data_1980s$nonunion_wage_2020_adj)

@@

### 2.4.2 Linear regression of percent collective bargaining and party state average

```{r}
decade_region <- list(
  c("1980s", "Northeast"),
  c("1980s", "West"),
  c("1980s", "North Central"),
  c("1980s", "South"),
  c("2010s", "Northeast"),
  c("2010s", "West"),
  c("2010s", "North Central"),
  c("2010s", "South")
)

@@

#### 2.4.2.1 Graph of regression estimate by decade and year

```{r}
Regression_plot <- ggplot(CB_party_regression, aes(x = Decade, y = Estimate, fill = Region)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = sprintf("P=%.3f", P_Value)),
             position = position_dodge(width = 0.9),
             vjust = -0.2,
             size = 3) +
  theme_minimal() +
  labs(title = "Linear Regression Estimate by Decade",
       subtitle = "Model of Percent Collective Bargaining vs Proportion of House Delegation That is Democratic")

@@
